---
output:
  word_document: default
  html_document: default
  pdf_document: default
---


Project: Detecting Heart Disease
Goal: to determine Whether a person suffers from heart disease using KNN 
Questions: 
   #• How well can we predict whether a patient suffers from heart disease or not
      #based on the predictor variables available to us?
   #• What value of k provides us with the best predictive performance? 
   #  (going to us knn)

----------------------------------------------------------------
 Load the data 
----------------------------------------------------------------
```{r}
library(dplyr)
library(stringr)
library(readr)

heart = "C:/Users/itzal/OneDrive/Desktop/UH LIFE/YEAR 4/ML R/ch 6/Student/Data/heart.csv"        
heart_data <- 
  read_csv(file=heart, col_types = "nffnnffnfnfnff")
glimpse(heart_data) # tells you all the features 
summary(heart_data) # tells you the summary of all numerical and categorical features 
```

----------------------------------------------------------------
 Data Cleaning: 
----------------------------------------------------------------
 • Dealing with missing data: 
    There's missing data for 10 of our 14 variables
    We are going to which is to remove the records with missing data from the dataset
```{r}
heart_data <- heart_data %>%
  filter(!is.na(restingBP) & !is.na(cholesterol) & !is.na(highBloodSugar) &
           !is.na(restingECG) & !is.na(restingHR) & !is.na(exerciseAngina) & 
           !is.na(STdepression) & !is.na(STslope) & !is.na(coloredVessels) & 
           !is.na(defectType))
 
# • Normalizing the Data using min-max normalization approach 
normalize <- function(x) {
   return((x - min(x)) / (max(x) - min(x))) 
}
#   normalization function to each of our numeric features to normalize their values 
#   between the range of 0 to 1.

heart_data <- heart_data %>%
  mutate(age = normalize(age)) %>%
  mutate(restingBP = normalize(restingBP)) %>%
  mutate(cholesterol = normalize(cholesterol)) %>%
  mutate(restingHR = normalize(restingHR)) %>%
  mutate(STdepression = normalize(STdepression)) %>%
  mutate(coloredVessels = normalize(coloredVessels))

summary(heart_data) # now range of values for  numeric features are all now within 0 and 1

# • Dealing with Categorical Features by creating dummy variables. 
#   A new dummy variable representing each of the unique values of the original 
#   categorical variable (ex sex variable --> sex_male, sex_female by using dummy variables )
heart_data <- data.frame(heart_data)

heart_labels <- heart_data %>% select(heartDisease)  #splitting of class labels from rest of the data as we do want to create a dummy variable for this  
heart_data <- heart_data %>% select(-heartDisease)

colnames(heart_data) # orignal features 

library(dummies)
heart_data <- dummy.data.frame(data=heart_data, sep="_")
colnames(heart_data) # now we have 25 features with 19 being dummy variables 


# • Splitting the Data: (75% train, 25% test)
set.seed(41)
sample_index <- sample(nrow(heart_data), round(nrow(heart_data)*.75), replace = FALSE)
heart_train <- heart_data[sample_index, ]
heart_test <- heart_data[-sample_index, ]

#  same split for class labels 
heart_train_labels <- as.factor(heart_labels[sample_index, ])
heart_test_labels <- as.factor(heart_labels[-sample_index, ])

# • Classifying Unlabeled Data 
#   Now ready to label our unlabeled data using the k-nearest neighbors approach
library(class)
heart_pred1 <-
  knn(
    train = heart_train,
    test = heart_test,
    cl = heart_train_labels,
    k = 15
  )
head(heart_pred1)
```

----------------------------------------------------------------
 Evaluating the Model 
----------------------------------------------------------------
  • need to compare the predicted labels for our test data (heart_pred1) 
   against the actual labels for our test data (heart_test_labels)
```{r}
heart_pred1_table <- table(heart_test_labels, heart_pred1)
heart_pred1_table
sum(diag(heart_pred1_table)) / nrow(heart_test)
# 81.33% accuracy
```
----------------------------------------------------------------
 Improving the Model by trying to modify the k value 
----------------------------------------------------------------
```{r}
heart_pred2 <-
  knn(
    train = heart_train,
    test = heart_test,
    cl = heart_train_labels,
    k = 1
  )
heart_pred2_table <- table(heart_test_labels, heart_pred2)
heart_pred2_table
sum(diag(heart_pred2_table)) / nrow(heart_test) #75% accuracy 

heart_pred3 <-
  knn(
    train = heart_train,
    test = heart_test,
    cl = heart_train_labels,
    k = 40
  )
heart_pred3_table <- table(heart_test_labels, heart_pred3)
heart_pred3_table
sum(diag(heart_pred3_table)) / nrow(heart_test) #84% accuracy 
```
